# Regression and model validation

*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

```{r}
date()
```
# the token to fix the connection

# usethis::create_github_token()
gitcreds::gitcreds_set()

# read the file

```{r}
list.files()
student2014 <- read.csv("data2.csv", header = TRUE, sep = ",")
dim(student2014)
str(student2014)
```
 
# Data is from an international survey of Approaches to Learning, made possible by Teachers and is  used as a course material in
# a course calles Introduction  to open data science. 
# dataset has seven variables two of which are integers and four of which are numerical variables. One variable describing gender is categorical. 

# variable contets briefly: 
# gender = gender of the respondent
# age = age of the respondent
# attitude = Global attitude toward statistics/10
# deep = mean of the values related to deep learning methods
# stra = mean of the values related to strategic learning methods
# surf = mean of the values related to surface learning methods
# points = exam points

# Graphic overview

```{r}
library("ggplot2")
library("GGally")

ggpairs(student2014, lower = list(combo = wrap("facethist", bins = 20)))
```

