---
title: "Chapter3.rmd"
output: html_document
date: "2022-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Logistic regression
## Diary
This week I learned more about logistic regression. This is probably one of the most useful parts of the course for me since I am going to use these methods on my seed bank germination data. Thus, I will likely return to these exercises and methods many times in the near future since this time I couldn't use as much time on these exercises as I would have wanted. This week I did not have any problems with committing or pushing, yay! 

## Reading the data and exploring the structure

````{r}
getwd()
setwd("/home/local/pielaura/Documents/zasemaemulaattori/Documents/kurssit/Introduction_to_open_data_science/excerices_1/IODS-project/data")
studentalc <- read.csv("alc.csv", sep = ",")
```

## The names of the variables and a brief description of the data

```{r}
colnames(studentalc)
dim(studentalc)
```
Data consists of questionnaires about alcohol consumption of secondary school students in Portugal. Data has variables concerning grades, demography, social features and school features. Dataset is a combination of two datasets addressing the effect of these variables on mathematics and Portuguese. There are 370 observations and 35 variables, 

## Relationship between high and low alcohol consumption on a selection of variables
I am investigating whether age, absence, sex or higher have an effect on whether the student consumes high or low amounts of alcohol. 

* Age - student's age (numeric: from 15 to 22) 
* Absence = number of school absences (numeric: from 0 to 93) 
* Sex = student's sex (binary: 'F' - female or 'M' - male) 
* Higher = wants to take higher education (binary: yes or no)

### Hypotheses
* age is not expected to have an effect on the alcohol consumption
* number of school absences is higher when the level of alcohol consumption increases
* males are predicted to have higher alcohol consumption than females. 
* students that want to take higher education use less alcohol than students that do not want to do so. 

## Data exploration

Firstly let's look at a bar plot made from each focal variable to see how the data looks like
```{r}
library(tidyr); library(dplyr); library(ggplot2)
focals <- select(studentalc, age, high_use, absences, sex, higher)
glimpse(focals)
focals$higher <- as.factor(focals$higher)
focals$sex <- as.factor(focals$sex)
gather(focals) %>% glimpse
gather(focals) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()
```


Then let's make a barplot showing the relationship between alcohol consumption and the focal variables
```{r}

g2 <- ggplot(focals, aes(x = high_use, y = absences, col = sex))
g2 + geom_boxplot() + ylab("absences") + ggtitle("Student's absences  by alcohol consumption and sex")

```


Seems like the students that have more absences quite often use more alcohol. The difference is more clear in males.

```{r}
g3 <- ggplot(focals, aes(x = high_use, y = age))

# define the plot as a boxplot and draw it
g3 <- g3 +  geom_boxplot() + ylab("age")

# initialize a plot of high_use and age
g3 <- ggplot(focals, aes(x = high_use, y = age, col = sex))

# define the plot as a box plot and draw it
g3 + geom_boxplot() + ylab("age") + ggtitle("Student's age by alcohol consumption and sex")

```

Older males seem to be using more alcohol than younger ones. For females there is no clear difference except in the average age, which seems to be lower in high alcohol-usage group. 

Next I will make cross tabulations of the binary variables that are age, higher and sex

```{r}
table(high_use = focals$high_use, sex = focals$sex)
table(high_use = focals$high_use, sex = focals$sex, higher = focals$higher)

```
Although there are less males in the data, there are still more males that use alcohol in high amounts. There are very few students that did not plan to go for higher education, thus a meaningful comparison between the differences is questionable. 

## Logistic regression

Relationship between the chosen variables and the high/low alcohol consumption. The chosen variables are age, absence, sex, higher

```{r}

m1 <- glm(high_use ~ age + absences + sex + higher, data = studentalc, family = "binomial")
# summary of the model 
summary(m1)
# coefficients
co <- coef(m1)
# coefficients as odds ratios
or <- coef(m1) %>% exp

# confidence intervals
ci <- confint(m1) %>% exp

# odds ratios with confidence intervals
cbind(or, ci)



```

Number of absences and male gender had significant effects on the probability of using high amounts of alcohol just as was hypothesized. Age did not have a significant effect on alcohol use, which was also expected. Plans on getting a higher education did not have an effect on the probability of a student to use high amounts of alcohol. However, there were only 15 students that did not plan on getting a higher education. Males were almost three times more likely using high amounts of alcohol than females. Also the as number of absences increased by one, increased the odds that the student uses high amounts of alcohol by 1.09. 

## Predictive power of my model

```{r}
# variables wit a statistical relationship with alcohol consumption
m2 <- glm(high_use ~ absences + sex, data = studentalc, family = "binomial")

# probability of high use (prediction)
probabilities <- predict(m2, type = "response")

# predicted probabilities to the original data
studentalc <- mutate(studentalc, probability = probabilities)

# prediction of high use
studentalc <- mutate(studentalc, prediction = probability > 0.5)

# how it looks like? 
select(studentalc, absences, sex, high_use, probability, prediction) %>% tail(10)


# tabulate the target variable versus the predictions
table(high_use = studentalc$high_use, prediction = studentalc$prediction)

```

```{r}

# plot probability vs. prediction
g <- ggplot(studentalc, aes(x = probability, y = high_use, col = prediction))+ggtitle("Probability vs. prediction") + geom_point()
g

# tabulate the target variable versus the predictions
table(high_use = studentalc$high_use, prediction = studentalc$prediction) %>% prop.table() %>% addmargins()

```
Training error (total proportion of inaccurately classified individuals)
0.23243243 + 0.02162162
= 0.254054

